### 2009-09-04 trunk/latex/paper1/contents.tex

@@

``softwareness''

Our hypothesis is that at least one of the presented models can synthesize
networks that resemble software networks. We already know that they produce
networks that, just like software networks, are scale-free. This single
property, though, isn't enough to prove the hypothesis, since there are many
known scale-free networks extracted from many distinct domains of knowledge. It
is important to distinguish between software and non-software networks.

@@

For this experiment we have collected 65 software systems written in the Java
programming language and 66 networks from other domains, such as biology,
sociology and linguistics. Both the software systems and the networks are freely
available on the Internet.

As of this writing, Java is a popular programming language, for which there is
plenty of open source software CITE. Also, there are many dependency extractors
for this language. We have extracted the dependency network for each software
system using a tool called Dependency Finder CITE, for its easy integration with
scripts. Dependency Finder looks for dependencies by means of static analysis of
Java bytecode and outputs the dependency network in a XML file. Although this
network is in a detailed level, showing dependencies between fine-grained
implementation entities such as attributes and methods, we have abstracted it to
a network of dependencies between classes. The network is directed and
non-weighted. The size of the networks range from XXX to YYY.

@@

The networks from other domains were collected in websites of complex network
research groups. The domains include sociology, linguistics, biology and
technology. The size of the networks range from XXX to YYY vertices.

@@

In assessing the realism of a model, we want to know if, with a proper choice of
parameters, the model is capable of producing a realistic network. In order to
do so, we must generate many networks using different combinations of
parameters.

Since most of the parameters can assume infinite distinct values, we chose to
fix some of them and vary the others in discrete steps. In all models the number
of vertices was fixed to 1000. We generated one network for each set of
parameters. Here we describe the criteria we use to choose the parameters for
each model.

@@

We chose five different module dependency networks, which where extracted from
actual dependencies between modules of five different software systems of our
sample, ranging from 2 to 32 modules. The module dependency networks are shown
on Figure \ref{fig:architectures}.

@@

The probabilities p, q and r were given all possible values from 0.0 upto 1.0,
in 0.2 steps, such as the sum of the probabilities was 1. Since the only events
that create vertices are those associated with probabilities p and q, we imposed
the additional restriction that p + q > 0.

@@

In a recent work, Milo et al. proposed to distinguish networks of different
domains by analyzing the triad concentration of each network. A triad is simply
a network with three vertices in which all vertices are connected. There are
only 13 distinct triads -- one for each configuration of directed edges --, as
shown in Figure \ref{fig:triads}.
\cite{Milo2002}
%CONSIDERING all the possible configurations of
%directed edges between three vertices, there are to 13 triads in which all
%vertices are connected, as shown in Figure X.

In order to characterize a network, one can count how many times each of the
13 triads appear on the network. The result is a triad concentration profile.
Figure 1a show triads for a software system... Figure 2a show triads for network
from domain X. \ref{fig:profiles}

Then, in order to measure how similar two networks are, Milo et al computed
Pearson's correlation coefficient between the two corresponding profiles.

CITE igraph

>>>>

\section{Characterization of Software Networks}

Our research hypothesis is that at least one of the presented models can
synthesize networks that resemble software networks. A central issue, thus, is
how to measure similarity between networks. In order to be useful, the
metric must yield high similarity between software networks and low similarity
between networks from distinct domains. In this section we present such a
metric, together with an experiment that evaluates the metric by applying it to
both software and non-software networks.

\subsection{A Metric for the Similarity Between Networks}

In a recent work, Milo et al. \cite{Milo2002} proposed to characterize networks
by analyzing their triad concentration. A triad is a network with three vertices
in which all vertices are connected. There are only 13 distinct triads -- one
for each configuration of directed edges --, as shown in Figure \ref{fig:triads}.

By counting how many times each triad appears in a network, one can build a
triad concentration profile (TCP). A TCP is, therefore, a vector with 13
elements that summarize the network. Figure \ref{fig:profiles} shows the TCP for
three distinct networks.

Similarity between two networks can be measured by computing the Pearson's
correlation coefficient between the corresponding TCPs, which yields a value
between -1 (most dissimilar) and 1 (most similar). This is the similarity metric
used in this research.

\subsection{Data Set}

To support the evaluation of the metric, we have collected 131 networks. The
networks are described in detail in Appendix A.

Software networks. We have collected 65 software systems written in Java. Java
is a popular programming language, for which there is plenty of open source
software.  The corresponding software networks were extracted by the tool
Dependency Finder CITE, considering classes and interfaces as components.
Dependency Finder was used because it can be easily integrated with scripts.

Non-software networks. We have collected 66 networks from distinct domains, such
as biology, sociology, technology, and linguistics. These networks are freely
available on the Internet and have been used in previous researches.

\subsection{Evalution of the Similarity Metric}

We measured the similarity between each pair of software networks and then
calculated the average and the standard deviation. The average similarity was
0.97 and the standard deviation, 0.03. The high similarity and the low standard
deviation shows that the metric indeed captures the commonalities in software
networks.

Then, for each non-software network, we measured the average similarity between
the network and all software networks.

In order to evaluate the similarity metric, we conducted an experiment on the
data set. ...
- high similarity between software networks;
- lower similarity between software networks and networks from other domains.

@@

Our research hypothesis is that at least one of the presented models can
synthesize networks that resemble software networks. In order to investigate the
hypothesis we conducted an experiment. First we generated many synthetic
networks, using all three models and different combinations of parameters.
After that we selected the networks that were more similar to software networks.
Finally, we sought patterns in the parameters of the selected networks.

A central issue is what

We assume that software networks possess a intrinsic quality, that we will call
``softwareness'', that distinguish them from networks in other domains. A
measure of the softwareness is essential to the experiment. One possible measure
is described and evaluated in the following sections. The measured softwareness
of a network represents how likely it is a software network.

% TODO: move to the section on the oracle
We already know that they
produce networks that, just like software networks, are scale-free. This single
property, though, isn't enough to prove the hypothesis, since there are many
known scale-free networks extracted from many distinct domains of knowledge.

To support the analysis of softwareness

@@

%We collected 65 software systems written in the Java
%programming language and 66 networks from other domains, such as biology,
%sociology and linguistics. Both the software systems and the networks are freely
%available on the Internet.
%
%As of this writing, Java is a popular programming language, for which there is
%plenty of open source software CITE. Also, there are many dependency extractors
%for this language. We have extracted the dependency network for each software
%system using a tool called Dependency Finder CITE, for its easy integration with
%scripts. Dependency Finder looks for dependencies by means of static analysis of
%Java bytecode and outputs the dependency network in a XML file. Although this
%network is in a detailed level, showing dependencies between fine-grained
%implementation entities such as attributes and methods, we have abstracted it to
%a network of dependencies between classes. The network is directed and
%non-weighted. The size of the networks range from XXX to YYY.

@@

%The networks from other domains were collected in websites of complex network
%research groups. The domains include sociology, linguistics, biology and
%technology. The size of the networks range from XXX to YYY vertices.

@@

We want to investigate if, with a proper choice of parameters, a model is
capable of synthesizing a network that resembles software networks. In this
section we describe which combinations of parameters values were used for each
model.

%In assessing the realism of a model, we want to know if, with a proper choice of
%parameters, the model is capable of producing a realistic network. In order to
%do so, we must generate many networks using different combinations of
%parameters.

In order to keep the networks in a tractable number, we have set the number of
vertices in each network to 1000, and varied the remaining parameters in
discrete steps such as each parameter assumed no more than 6 values. We have
generated one network for each combination of parameter values.

%Since most parameters are continuous, we had to 
%we chose to threat the number of vertices
%as a constant and vary the other parameters in discrete steps. In all models the
%number of vertices was fixed to 1000. We generated one network for each set of
%parameters.
%Here we describe the criteria we use to choose the parameters for each model.

@@

We chose five different module dependency networks, which where extracted from
actual dependencies between archives of five different software systems of our
sample. The module dependency networks are shown on Figure
\ref{fig:architectures}.

@@

The probabilities p, q and r were given all possible values from 0.0 up to 1.0,
in 0.2 steps, such as the sum of the probabilities was 1. Since the only events
that create vertices are those associated with probabilities p and q, we imposed
the additional restriction that p + q > 0.

@@
